---
title: "removing_reflections"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{removing_reflections}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Three methods for removing reflections

- 'Next reasonable speed' method
- 'Next reasonable distance' method
- 'Next extreme speed' method

## Load libraries

```{r warning=FALSE, message=FALSE}
library(tidyr)
library(dplyr)
library(ggplot2)
```

## Next reasonable speed method

This method fails for any reflections of duration $T$, where $T = D/v$, where $D$ is the displacement of the reflected points from the anchor point, and $v$ is the speed cutoff used.

```{r}
# get displacement distance and speed cutoff combinations
displacement_distance <- c(10 ^ seq(0, 4, 0.1))
speed_cutoffs <- seq(0, 100, 5)

# use formula for algorithm defeat
data <- crossing(displacement_distance, speed_cutoffs) %>% 
  mutate(time_to_defeat = displacement_distance / speed_cutoffs)
```

```{r fig.cap="Figure showing the time (in minutes) required to beat the 'next reasonable speed' method, for various combinations of the reflection distance from the real positions, _D_ and the speed cutoff, _v_. Area in black represent combinations that fail"}
# plot matrix of time to defeat
ggplot(data)+
  geom_tile(aes(speed_cutoffs, displacement_distance,
                fill = time_to_defeat / 60))+
  scale_fill_viridis_c(option = "C",
                       trans = "log10",
                       limits = c(0.2, NA),
                       breaks = c(0.3, 1, 3, 10, 30),
                       labels = as.character(c(0.3, 1, 3, 10, 30)),
                       na.value = "black")+
  scale_y_log10(labels = scales::comma,
                breaks = c(10 ^ seq(0, 5), 500, 5000))+
  scale_x_continuous(breaks = c(10, seq(20, 100, 20)))+
  coord_equal(ratio = 50,
              xlim = c(10, NA),
              ylim = c(100, NA))+
  theme_minimal()+
  theme(axis.text.y = element_text(angle = 90, 
                                   hjust = 0.5),
        legend.position = "top")+
  labs(x = expression("speed cutoff" ~ italic("v") ~ "(m/s)"),
       y = expression("displacement distance" ~ italic("D") ~ "(m)"),
       fill = "reflection duration (minutes)")
```

## Next reasonable distance method

```{r fig.cap="An example of a reflection, indicated by the dark line away from the real track shown in a lighter colour."}
# simulate some data of length 1000
library(smoove)
nu <- 2
tau <- 5
dt <- .1 
ucvm1 <- simulateUCVM(nu=nu, tau=tau, T.max = 100, dt = dt)

# add a reflection to x_max + x_max / 10, y_max + y_max / 10
data <- as_tibble(ucvm1$XY) %>% 
  mutate(time = seq_len(nrow(.)))
data <- data[seq_len(1000), ]
data[(nrow(data) / 2):(nrow(data) / 2) + 50, "x"] <- max(data$x) + (max(data$x) / 10)
data[(nrow(data) / 2):(nrow(data) / 2) + 50, "y"] <- max(data$y) + (max(data$y) / 10)

# get speed estimate
data$speed <- atlastools::atl_get_speed(data)

ggplot(data)+
  geom_point(aes(x, y,
                col = speed),
             shape = 16, size = 0.1)+
  geom_path(aes(x, y,
                col = speed),
            size = 0.1)+
  scale_colour_viridis_c(direction = -1,
                         option = "E")+
  theme_void()+
  labs(x = NULL, y = NULL)
```

WIP.
