---
title: "Iterating patch construction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Iterating patch construction}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

atlastools was built to handle large tracking datasets from the [WATLAS system](https://www.nioz.nl/en/about/cos/coastal-movement-ecology/shorebird-tracking/watlas-tracking-regional-movements). 
These datasets are -- currently -- split into subsets and saved locally, typically in a folder called _revisitData, revisit_data_ or similar [^1]. In general, atlastools packages would work on the whole dataset as well, but this functionality has not been tested.

Here, we cover iterating atlastools functions over multiple data files, and introduce the `wat_repair_ht_patches` function.

```{r, include = FALSE}
# knitr::opts_chunk$set(
#   collapse = TRUE,
#   comment = "#>"
# )
```

```{r setup, message=FALSE}
library(atlastools)

# read data and see leading lines
library(data.table)
library(purrr)
library(magrittr) # pipe operator %>%

# for plotting
library(ggplot2)
```

# Iterating over tidal cycles

Begin by reading in the data files from 10 tidal cycles for one individual.

```{r read_data, message=FALSE}
# list the data files available and read them in
data_files <- list.files("vignette_data/", pattern = "413", 
                         full.names = TRUE)

data <- map(data_files, fread)
```

Apply atlastools patch functions on each tidal cycle. atlastools functions are tidyverse compatible, and can be chained using the `%>%` operator. Here, we wil use the function defaults, but they can be passed as arguments.

```{r apply_atlastools, message=FALSE}
# map over the list of data.tables
patch_data <- map(data, function(df){
  to_return <- df %>% 
    wat_infer_residence() %>% 
    wat_classify_points() %>% 
    wat_make_res_patch()
  return(to_return)
})
```

The data returned is a list; each element of dataframes (really of the class `data.table`). 
The dataframes in the list contain a list-column of dataframes (a dataframe inside another dataframe); these nested dataframes hold the point data that was used to construct each patch.

# Repairing high tide patches

High tide presents an issue for atlastools, since it was originally written with low tide data in mind. The issue is that residence patches at the beginning of a tidal cycle (high tide) are essentially continuations of residence patches at the end of the previous tidal cycle.

The `wat_repair_ht_patches` function was written to fix this issue. It is a reworking of the original `wat_make_res_patch` function, except that it operates on a list of residence patches, as we made above, and looks only at the first and last patches of each tidal cycle.

If the first patch (_Pa_) of tidal cycle (called `tidenumber` in the data) is not spatio-temporally distinct from the last patch of the previous tidal cycle (_Pb_), _Pa_ is merged into _Pb_.

```{r repair_ht_patches}
# count patches in unrepaired data
patch_data <- compact(patch_data) # remove NULL elements
map_int(patch_data, nrow) %>% sum()

# repair patches for id 413
repaired_data <- wat_repair_ht_patches(patch_data_list = patch_data)

# count the number of repaired patches
nrow(repaired_data)
```

From an initial unrepaired number of `r map_int(patch_data, nrow) %>% sum()`, we now have `r nrow(repaired_data)` patches.

# Visualising repaired data

Here, we will visualise the repaired data as in the previous vignette.

```{r vis_rep_data}
# plotting libraries
library(ggplot2)

fig_patch_repair <- ggplot(repaired_data, aes(x_mean, y_mean))+
  geom_path(aes(group = tide_number),
            col = "black", lwd = 0.2, lty = 2)+
  geom_point(aes(size = duration/60,
                 col = factor(tide_number)),
             stroke = 1,
             shape = 21, fill = NA,
             alpha = 0.9)+
  geom_text(aes(label = patch), 
            col = "grey20", size = 2)+
  
  scale_size(range = c(3,10))+
  scale_colour_brewer(palette = "Paired", direction = 1)+
  
  scale_x_continuous(expand = expansion(mult = c(0.1)))+
  scale_y_continuous(expand = expansion(mult = c(0.1)))+
    
  coord_sf(crs = 32631)+
  # annotation_scale(location = "br", width_hint = 0.1, text_cex = 1)+
  
  theme_bw()+
  theme(legend.position = "bottom",
        axis.text = element_blank(),
        axis.ticks = element_blank())+
  
  labs(x = NULL, y = NULL, 
       col = "# tide", size = "residence (mins)")
```

```{r show_patches, fig.width=8, fig.height=6}
fig_patch_repair
```

